### 2023-06-23 chapter_2/src/appendix.qmd
study 2 models done for CH2git add -A! Just need to write

>>>>

## Hypothesis 5

For $H_5$, I am interested in examining whether or not districts that lean towards one party more than the other have candidates use more of the color in line with the party that has a historical advantage and less of the color for the party that tends to suffer electorally speaking.

As I mention in the main text of the book, I note that I regress the 5-year rolling average of the democratic vote share in a state upon the proportion of the color red and blue on the yard signs downloaded from the Center for American Political Design. The outcome is a continous and bounded variable between the value of 0 and 1. I note in the main text that I use a ordered beta regression, but here I provide more specifics. As the outcome is continuous but bounded between 0 and 1, treating the outcome variable as a continous variable with no finite bounding as I would with a OLS would be inappropriate. To ensure that I am using the most appropriate estimator to explain the outcome of interest, I elect to use an ordered beta regression. Beta regression models are a type of distributional model that essentially tries to model the distribution of the outcome variable. The standard beta regression model can fit models where the outcome is (0, 1). However, my outcome is [0, 1). I could elect to use a zero-inflated beta regression which would allow me to appropriately model this outcome. However, @kubinec_2022_pa demonstrates that the ordered beta regression is a much more efficient and accurate (in terms of a number of various metrics) approach to dealing with outcomes that can equal either 0 or 1.

The default priors for $\beta_i$ is $N(0, 5)$ in the $ordbetareg$ [@kubinec_2022_pa] package that I use to fit the model. However, I receive warnings indicating that there are divergent transitions. For HMC, divergent transitions indicate that the sampler is having problems with the curvature of your posterior distribution. As I note in the main text of the book, I am using a multilevel model that includes a partially pooled parameter for the districts and use two non-pooled parameters for the state and the election year. This is a pretty complex problem. Especially givent the estimator that I am using here. More troubling still is that my $\hat{R}$ values are not equal to 1 for either model. These two things together suggest that there is a pretty serious problem and that my posterior distributions are quite unstable.

To try to fix this issue, I adjust the prior for $\beta_1$ to $N(0, 2)$. While I still receive a warning about divergences in both models, the number of them drops down to about 30 out of 4000. Additionally the $\hat{R}$ values are at 1. Together, these tell me that these priors likely helped the models quite a bit and feel more confident in the stability of my model. @tbl-hypothesis-5 presents the results of this second specification of the model.

```{r}
#| label: tbl-hypothesis-5
#| tbl-cap: The effect of democratic vote share on color choices for yard sign

names(list_fitted[["h_5"]]) <- c("Prop. yard sign Red", "Prop. yard sign Blue")
modelsummary(
  list_fitted[["h_5"]]
  , stat = "conf.int"
  , conf_level = 0.9
  , metrics = c("n.obs", "RMSE")
  , coef_map = c(
    "b_intercept" = "Intercept"
    , "b_dem_vote_share" = "Democratic vote share (%)"
  )
)

```

### 2023-06-23 chapter_2/src/manuscript.qmd
study 2 models done for CH2git add -A! Just need to write

>>>>

[Study 1](#sec-study-1) presents evidence in support of the idea that the colors red and blue matter to politics -- they convey information to people. Second, the evidence from [Study 1](#sec-study-1) suggests that the information can shape the attitudes of voters. What remains unclear is whether or not these patterns have large enough impacts that it shapes how campaigns market themselves. In informal conversations with a campaign marketing firm which specializes in campaign marketing, I was told:

> There obviously is a strong association with blue and the Democratic Party and red and the Republican Party.  We've seen in some of our testing that specifically MAGA red tends to elicit a slightly stronger association with Republicans/conservatives/Trump than other shades. 
> The rest of color spectrum doesn't seem to have as strong an association, though greens and purples tend to read a little more left, the combination of red, white and blue tends to read a little more right, black and white tends to read a little left, black and red tend to read a little more right etc.

While this was good information to hear that seemed to align with what I found in [Study 1](#sec-study-1), I wasn't sure how much this actually influenced strategy. It appears it does, however.

> Beyond that,  I generally suggest that considering associations with party is generally a good place to start.  For example, if you're running in a Democratic primary and that's the race that's going to most likely decide who ultimately gets elected, then blues are a good place to start.

While there is some qualification in that there are other colors that can be useful beyond red and blue. The value of colors associated with a particular place can be a way for candidates to communicate attachment to the district and to the constituents. Though, there is a preference for campaigns to use as few colors as possible to save on costs - simpler is better. What particularly stood out to me was when I was told

> You want voters to be able to look at whatever collateral you have, including lawn signs, and be able to identify that it's your material without ever having to read it. If you do one thing long enough, voters tend to start to associate specific color schemes with specific candidates.

So, unsurprisingly, campaigns care about how they market themselves. And it seems as though campaigns want to tap into quick and efficient information processing for voters and potential voters. However, as noted during the conversation

> A lot of decisions are still made on purely gut feelings as opposed to taking a more multi-disciplinary evidence-based approach.

The goal for the second peice of empirical evidence in the chapter is to explore whether there is any evidence not from the "lab" suggesting that campaigns use the colors red and blue systematically. If the color of a yard sign does indeed matter and its electoral effects depend on the perceptions of potential voters, we should expect that campaigns are less likely to place yard signs with a large amount of Red or Blue in more moderate districts. However, in districts where there is a strong partisan base, we should expect that the candidate of the popular party will use the party's color more, while the disadvantaged candidate will use their party's color less as a way to either play up or down their loyalty to their party.

First, I collect data from the MIT Election Lab that reports the electoral results of House of Representative elections from the years 1970-2020. The data here tell me which candidate got how many votes in each House election across the United States. With this data, I calculate the rolling 5-year average of the Democratic party's vote share for each district.

Next, I download around 900 yard signs from House elections from the years 2016 to 2024 from the website provided by the Center for American Politics and Design. I use the `OpenCV` library from python to examine the proportion of the colors near the "Republican Red"^[lower values: 93, 9, 12; upper values 236, 69, 75] and "Democratic Blue"^[lower values: 0, 18, 26; upper values: 102, 212, 255] in each of these yard signs.

For example, I downloaded the GOP logo used on their official Twitter account during the 2022 midterm election cycle. What `OpenCV` does is it loads the image and converts it to a three-dimensional `numpy` array that contains the GBR (reverse of RBG) values for the pixels in the image. I standardize the image to be 224 $\times$ 224 pixels. The color is trained to detect the range of Republican Red and Democrat Blue colors. Once trained, the computer detects the pixels that do not contain values within the pre-specified range and converts those pixels to be black.

```{python}
#| label: example detection
#| echo: false

from sys import path # for path management
from cv2 import imread, imwrite # for color detection
    #* user-defined
#path.append("../")
from PY.helper import colorDetector

# Define colors to detect
    #* White
        #** Not defined. Default for colorDetector()
    #* Red
republican_red = [232, 27, 35] # target color
red_lower = [93, 9, 12] # lower end of spectrum for red
red_higher = [237, 69, 75] # higher end of spectrum for red

# Load gop image to read
img = imread("./assets/img/gop_2022.png")

# Detect colors
percent, img_transformed, result = colorDetector(img = img, color_upper = red_higher, color_lower = red_lower)

transformed= imwrite("../data/temp/study_2/gop_2022_transformed.png", img_transformed) # save transformed image

masked= imwrite("../data/temp/study_2/gop_2022_detected.png", result) # save masked image
```

::: {#fig-color-detection-example layout-ncol="2"}
![Resized original image](../data/temp/study_2/gop_2022_transformed.png){#fig-resized}

![Masked](../data/temp/study_2/gop_2022_detected.png){#fig-masked}

Detecting colors in the GOP logo
:::

I then extract the non-black pixels in the array and calculate the percentage of pixels in the image that are not black.^[$\text{Color} \% = \frac{\text{Non-black}}{\text{Transformed}} \times \frac{\text{Original}_{\text{Height}} + \text{Original}_{\text{Width}}}{2\text{Transformed}_{\text{Height}} + 2\text{Transformed}_{\text{Width}}}$]

```{python}
#| output: asis
# Print calculated percentage of example
print("For the example in @fig-color-detection-example, about {:.2f}".format(percent) + " of the image is red.")
```

With the data on the proportion of each of the yard signs using the red and blue colors, I merge this data with the MIT Election Lab Data. This leaves me with data from the House elections between 2016 and 2020. While the data are limited in the number of elections it reflects, it allows me to hold the district composition constant as redistricting occurs after the Census every 10 years. However, each election cycle is different and are often different depending on current events. With these data, I estimate a mixed effect model with partial pooling of the intercepts at the district level and no pooling of the intercepts for the election year level. For less technical readers, this allows me to explore whether the district and within election year differences between choices in yard sign color use. What I am trying to do with this analysis is explore whether or not campaigns in districts that are more partisan will use more red or blue while campaigns in districts that are more moderate will use fewer of those traditionally partisan colors.

```{r}
#| label: fig-study-2-pred
#| layout-ncol: 1
#| fig-cap: More blue on the yard signs in Democratic strongholds
#| fig-subcap:
#|   - Proportion of red
#|   - Proportion of blue
red <- predicted_prob_bar(
  fitted_model = list_fitted[["h_5"]][["red"]]
  , treatment = "Red"
  , hypothesis = "H5"
)
blue <- predicted_prob_bar(
  fitted_model = list_fitted[["h_5"]][["blue"]]
  , treatment = "Blue"
  , hypothesis = "H5"
)
red
blue
```
